### 2015-04-06 cug_paper/03_analysis.tex
Paper: Grammar & punctuation editing for analysis & conclusion sections

For both test series, the daemon was configured to write each rank's data to a separate file and the test measured the application's perceived write performance as the write size increased.  Note the word 'perceived'.  What's being measured is how long it takes for each rank of the application to copy its data into the GPU memory.   The GPU cards in each of Titan's compute nodes have 6GB of RAM, though in practice only a little over 5GB is actually available to the user.  This means that for the first test series, with 8 ranks per node, write sizes of up to 512 MB were small enough for all ranks' writes to fit into the GPU memory.  For the second test series, using 16 ranks per node, write sizes up to 256MB would fit.  For both test series, once the write size exceeded the available GPU memory, the ranks would have to wait until the daemon was able to drain data out of the GPU memory and into the filesystem.

@@

Figure \ref{fig:results_8_nobars} shows the remaining results of the first test series.  This series was run with 8 ranks per node and with 1, 2 \& 4 write threads in the daemon.  The graph shows a number of interesting features.  The first and most obvious, is that multiple write threads decrease performance.  Exactly why this is so is unclear, but it appears that having multiple threads read from the GPU's memory interfered with the individual ranks' ability to write to the GPU's memory.

@@

Figures \ref{fig:results_base_16} and \ref{fig:results_16_nobars} show the results for the second series of tests.  As noted above, this series used 16 ranks per node, plus the daemon's threads.  This meant, of course, that the cores were oversubscribed.  Notice that Figure \ref{fig:results_base_16} as the same basic shape as Figure \ref{fig:results_base_8}.  The only significant difference is that the reported speeds shown in Figure \ref{fig:results_base_16} are approximately half those shown in Figure \ref{fig:results_base_8}.  This is expected, since there are twice as many ranks writing.

@@

Figures \ref{fig:results_base_8} and \ref{fig:results_base_16} show that Lustre client cache can provide good write performance up to 256MB write sizes.  However, that presumes that there is, in fact, sufficient system memory available on the node \emph{and} that the individual ranks are writing to different OST's.  If a 'real' application only wrote to a single file, the Lustre cache would not provide nearly as large a benefit unless the user took additional steps to increase the stripe count of the file.  It's worth noting that if an application used the techniques demonstrated in this paper, it would see improved write performance regardless of how many OST's were used and how much system ram was available for cache.

Looking at figure \ref{fig:results_16_nobars}, note the drop in performance between 16MB and 32MB and the slight downward trend from 32MB to 256MB.  This pattern also appears in Figure \ref{fig:results_8_nobars}, but is much more obvious in Figure\ref{fig:results_16_nobars}.  The authors hypothesize that this is related to the response time of the daemon.  As mentioned earlier, the daemon allocates GPU memory in blocks of up to 16MB.  Thus, if one of the ranks wants to write more than 16GB, it will need to go through more than one request/response message cycle.  However, the daemon only has a single thread to handle these messages.  When multiple ranks are sending message, they will necessarily have to wait on a response from the daemon and this slow their overall performance.  This is more noticeable in Figure\ref{fig:results_16_nobars} because there are more ranks making requests.
In short, it appears that improving the response time of the message handling thread would be beneficial.

>>>>

For both test series, the daemon was configured to write each rank's data to a separate file and the test measured the application's perceived write performance as the write size increased.  Note the word 'perceived'.  What was actually measured was how long it took for each rank of the application to copy its data into the GPU's memory.   The GPU cards in each of Titan's compute nodes have 6GB of RAM, though in practice only a little over 5GB is actually available to the user.  This means that for the first test series, with 8 ranks per node, write sizes of up to 512 MB were small enough for all ranks' writes to fit into the GPU memory.  For the second test series, using 16 ranks per node, write sizes up to 256MB would fit.  For both test series, once the write size exceeded the available GPU memory, the ranks would have to wait until the daemon was able to drain data out of the GPU memory and into the filesystem.

@@

Figure \ref{fig:results_8_nobars} shows the remaining results of the first test series.  As stated earlier, this series was run with 8 ranks per node and with 1, 2 \& 4 write threads in the daemon.  The graph shows a number of interesting features.  The first and most obvious, is that multiple write threads decrease performance.  Exactly why this is so is unclear, but it appears that having multiple threads read from the GPU's memory interfered with the individual ranks' ability to write to the GPU's memory.

@@

Figures \ref{fig:results_base_16} and \ref{fig:results_16_nobars} show the results for the second series of tests.  As noted above, this series used 16 ranks per node, plus the daemon's threads.  This meant, of course, that the cores were oversubscribed.  Notice that Figure \ref{fig:results_base_16} has the same basic shape as Figure \ref{fig:results_base_8}.  The only significant difference is that the reported speeds shown in Figure \ref{fig:results_base_16} are approximately half those shown in Figure \ref{fig:results_base_8}.  This is expected, since there are twice as many ranks writing.

@@

Looking at figure \ref{fig:results_16_nobars}, note the drop in performance between 16MB and 32MB and the slight downward trend from 32MB to 256MB.  This pattern also appears in Figure \ref{fig:results_8_nobars}, but is much more obvious in Figure 
\ref{fig:results_16_nobars}.  The authors hypothesize that this is related to the response time of the daemon.  As mentioned earlier, the daemon allocates GPU memory in blocks of up to 16MB.  Thus, if one of the ranks wants to write more than 16GB, it will need to go through more than one request/response message cycle.  However, the daemon only has a single thread to handle these messages.  When multiple ranks are sending message, they will necessarily have to wait while the daemon services any previous messages and this slows their overall performance.  This is more noticeable in Figure 
\ref{fig:results_16_nobars} because there are more ranks making requests.
In short, it appears that improving the response time of the message handling thread would be beneficial.

### 2015-04-06 cug_paper/05_conclusion.tex
Paper: Grammar & punctuation editing for analysis & conclusion sections

For the first test series (8 ranks/node), the results show clear performance improvement compared to writing directly to the filesystem.   More data can be cached in the GPU's memory than in the Lustre client-side cache and the average write speed is, surprisingly, slightly higher.  It is, however, necessary to use relatively large writes.  As shown in Figure \ref{fig:transfer_bw}, the transfer bandwidth between the host and GPU is relatively poor for small transfers.  Comparing Figure \ref{fig:results_base_8} and Figure \ref{fig:results_8_nobars},  it can be seen that write sizes must be at least 4MB before transfers to the GPU memory outperform writes to the Lustre cache.

@@

The results for the second test series (16 ranks/node) are less clear.  As with the first test series, write sizes need to be at least 4MB Looking closely at Figure \ref{fig:results_base_16} and Figure \ref{fig:results_16_nobars}, it appears that write sizes that can fit into GPU RAM perform slightly better than just writing to disk.  However, write sizes that exceed the capacity of the GPU RAM are actually a little slower than writing straight to the filesystem.

There are several other complications to consider, though.
First, Since the test application did no actual computation, there were plenty of cores available for the daemon to execute on.  In a real HPC application, this wouldn't be the case and the application's performance would almost certainly be negatively affected by the context switches.  On the other hand, an application that is running 16 ranks/node is unlikely to leave much system memory free for use by the Lustre client-side cache.  So, caching writes in the GPU memory might offset the disruption caused by oversubscribing the cores.  Whether or not this technique provides a net performance improvement in such a configuration will probably depend on the particular application.

@@

In general, the results show that using the GPU memory for cache provides approximately the same performance as the Lustre client-side cache.  The difference is that in order to benefit from the Lustre cache, the writes must be spread across a sufficient number of OST's and the memory must actually be available.  Whether or not these conditions can be met depends on the particular application.

>>>>

For the first test series (8 ranks/node), the results show a clear performance improvement compared to writing directly to the filesystem.   More data can be cached in the GPU's memory than in the Lustre client-side cache and the average write speed is, surprisingly, slightly higher.  It is, however, necessary to use relatively large writes.   Comparing Figure \ref{fig:results_base_8} and Figure \ref{fig:results_8_nobars},  it can be seen that write sizes must be at least 4MB before transfers to the GPU memory outperform writes to the Lustre cache.

@@

The results for the second test series (16 ranks/node) are less clear.  As with the first test series, write sizes need to be at least 4MB Looking closely at Figure \ref{fig:results_base_16} and Figure \ref{fig:results_16_nobars}, it appears that write sizes that can fit into the GPU's memory perform slightly better than just writing to disk.  However, write sizes that exceed the capacity of the GPU's memory are actually a little slower than writing straight to the filesystem.

There are several other complications to consider, though.
First, Since the test application did no actual computation, there were plenty of cores available for the daemon to execute on.  In a real HPC application, this wouldn't be the case and the application's performance would almost certainly be negatively affected by the context switches.  On the other hand, an application that is running 16 ranks/node is unlikely to leave much system memory free for use by the Lustre client-side cache.  So, the improvement from caching writes in the GPU's memory might offset the disruption caused by oversubscribing the cores.  Whether or not this technique provides a net performance improvement in such a configuration will probably depend on the particular application.

@@

In general, the results show that using the GPU's memory to cache filesystem writes provides approximately the same performance as the Lustre client-side cache.  The difference is that in order to benefit from the Lustre cache, the writes must be spread across a sufficient number of OST's and the memory must actually be available.  Whether or not these conditions can be met depends on the particular application.

%Figures \ref{fig:results_base_8} and \ref{fig:results_base_16} show that Lustre client cache can provide good write performance up to 256MB write sizes.  However, that presumes that there is, in fact, sufficient system memory available on the node \emph{and} that data is being written to a sufficient number OST's.  If a 'real' application only wrote to a single file, the Lustre cache would not provide nearly as large a benefit unless the user took additional steps to increase the stripe count of the file.  It's worth noting that if an application used the techniques demonstrated in this paper, it would see improved write performance regardless of how many OST's were used and how much system ram was available for cache.
